Using Occupancy Theory and Hierarchical Models to Estimate Otter Population in Southeastern Minnesota

### 2010-02-24 Paper.tex
Added rest of my stuff.
Added Kiva's citations.

    \subsection{WinBUGS/R}
    To be filled in...

\section{Standard Hierarchical Model}
Also to be filled in... :(.

@@

    \subsection{Conditional autoregressive (CAR) models}
    Let $\theta_s$ be the probability of a track being laid down in site $s$ on
    a given day. Then the Gaussian CAR model can be defined using the following
    conditional distributions:
    \begin{equation}
        \theta_s|\theta_i,\tau_s^2 \sim N(\mu_s+\sum_{i\in N_s}
        c_{si}(y_i-\mu_i),\tau_s^2)
    \end{equation}
    where $N_s$ is the set of neighbors of area $s$ (in this case, the two
    bordering sites), $E(\theta_s)=\mu_s$, $\tau_s^2$ is the conditional
    variance, $c_{si}$ is the weight defining how strong the correlation is
    between sites $s$ and $i$, and $c_{ii}=0$ for all sites $i$ (Arab et al.,
    year?). Essentially, this model means that if site $s$'s neighbor has a
    particularly high or low track-laying probability, that will increase or
    decrease the track-laying probability for site $s$. This makes it more
    likely to have strings of tracks that cross over multiple sites.

    \subsection{Our CAR model}
    \label{car model}
    For our model, both neighbors are weighted equally. Since $\tau_s$ for any
    site $s$ is unknown, we followed convention and used the same value of
    $\tau$ for all sites and placed a prior distribution on it. We chose a
    $\gamma(0.5,0.0005)$ prior distribution for $(\tau^2)^{-1}$, a common choice
    (GeoBUGS manual). In our model we define $\text{logit}(\theta_s)=\alpha_0+
    \alpha_{1_s}$ where $\alpha_0$ has an improper uniform distribution and
    $\alpha_1$ has a Gaussian CAR distribution. We again used a uniform prior
    distribution on $\psi$ and the same beta prior distributions on $p$ and $E$.
    In this case, we limited the range of $p$ to the interval $(0.7,1)$ and the
    range of $E$ to $(0,0.3)$. We increased the burn-in time to 12,000
    iterations. With this model, instead of a single estimate for $\theta$ there
    are now $S$ estimates for $\theta$ where $S$ is the total number of sites.
    While ideally we would estimate an overall value for $\theta$, it seems more
    important for the purposes of this investigation to estimate $\psi$.

>>>>

    \subsection{WinBUGS/R}
    All of our statistical analysis was performed in R.

    All of our modeling was performed using BRugs, which allowed us to use
    WinBUGS commands and other stuff in R.

\section{Standard Hierarchical Model}

    \subsection{Model Parameters}

    \subsection{Simulated Data}
    \noindent{\begin{tabular}{|l|l|l|l|}
        \hline
        \multicolumn{4}{|l|}{\textbf{The Model vs. Simulated Data}} \\
        \hline
            \(\psi\) & \(\theta\) & Convergence & 95\% CI (\(\psi\)) \\
        \hline
        \multirow{5}{*}{0.20}
            & \cellcolor{Green}0.20 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.112, 0.305) \\
            & \cellcolor{Green}0.40 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.142, 0.341) \\
            & \cellcolor{Green}0.60 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.098, 0.284) \\
            & \cellcolor{Green}0.80 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.097, 0.281) \\
            & \cellcolor{Yellow}1.00 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.117, 0.302) \\
        \hline
        \multirow{5}{*}{0.40}
            & \cellcolor{Yellow}0.20 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.388, 0.721) \\
            & \cellcolor{Green}0.40 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.270, 0.494) \\
            & \cellcolor{Green}0.60 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.287, 0.512) \\
            & \cellcolor{Red}0.80 & \cellcolor{Red}Bad &
              \cellcolor{Red}(0.277, 0.507) \\
            & \cellcolor{Yellow}1.00 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.279, 0.503) \\
        \hline
        \multirow{5}{*}{0.60}
            & \cellcolor{Yellow}0.20 & \cellcolor{Yellow}Good &
              \cellcolor{Yellow}(0.312, 0.581) \\
            & \cellcolor{Green}0.40 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.400, 0.630) \\
            & \cellcolor{Yellow}0.60 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.462, 0.687) \\
            & \cellcolor{Yellow}0.80 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.502, 0.727) \\
            & \cellcolor{Red}1.00 & \cellcolor{Red}Bad &
              \cellcolor{Red}(0.318, 0.549) \\
        \hline
        \multirow{5}{*}{0.80}
            & \cellcolor{Green}0.20 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.578, 0.838) \\
            & \cellcolor{Green}0.40 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.736, 0.915) \\
            & \cellcolor{Yellow}0.60 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.587, 0.829) \\
            & \cellcolor{Yellow}0.80 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.599, 0.811) \\
            & \cellcolor{Red}1.00 & \cellcolor{Red}Bad &
              \cellcolor{Red}(0.612, 0.830) \\
        \hline
        \multirow{5}{*}{1.00}
            & \cellcolor{Yellow}0.20 & \cellcolor{Yellow}Good &
              \cellcolor{Yellow}(0.895, 0.999) \\
            & \cellcolor{Yellow}0.40 & \cellcolor{Yellow}Okay &
              \cellcolor{Yellow}(0.927, 0.999) \\
            & \cellcolor{Green}0.60 & \cellcolor{Green}Good &
              \cellcolor{Green}(0.949, 1.000) \\
            & \cellcolor{Yellow}0.80 & \cellcolor{Yellow}Good &
              \cellcolor{Yellow}(0.802, 0.954) \\
            & \cellcolor{Red}1.00 & \cellcolor{Red}Bad &
              \cellcolor{Red}(0.800, 0.955) \\
        \hline
    \end{tabular}}

@@

    \subsection{Conditional autoregressive (CAR) models}
    Let $\theta_s$ be the probability of a track being laid down in site $s$ on
    a given day. Then the Gaussian CAR model can be defined using the following
    conditional distributions:
    \begin{equation}
        \theta_s|\theta_i,\tau_s^2 \sim N(\mu_s+\sum_{i\in N_s}
        c_{si}(y_i-\mu_i),\tau_s^2)
    \end{equation}
    where $N_s$ is the set of neighbors of area $s$ (in this case, the two
    bordering sites), $E(\theta_s)=\mu_s$, $\tau_s^2$ is the conditional
    variance, $c_{si}$ is the weight defining how strong the correlation is
    between sites $s$ and $i$, and $c_{ii}=0$ for all sites $i$ \cite{Arab2008}.
    Essentially, this model means that if site $s$'s neighbor has a
    particularly high or low track-laying probability, that will increase or
    decrease the track-laying probability for site $s$. This makes it more
    likely to have strings of tracks that cross over multiple sites.

    \subsection{Our CAR model}
    \label{car model}
    For our model, both neighbors are weighted equally. Since $\tau_s$ for any
    site $s$ is unknown, we followed convention and used the same value of
    $\tau$ for all sites and placed a prior distribution on it. We chose a
    $\gamma(0.5,0.0005)$ prior distribution for $(\tau^2)^{-1}$, a common choice
    \cite{Thomas2004}. In our model we define $\text{logit}(\theta_s)=\alpha_0+
    \alpha_{1_s}$ where $\alpha_0$ has an improper uniform distribution and
    $\alpha_1$ has a Gaussian CAR distribution. We again used a uniform prior
    distribution on $\psi$ and the same beta prior distributions on $p$ and $E$.
    In this case, we limited the range of $p$ to the interval $(0.7,1)$ and the
    range of $E$ to $(0,0.3)$. We increased the burn-in time to 12,000
    iterations. With this model, instead of a single estimate for $\theta$ there
    are now $S$ estimates for $\theta$ where $S$ is the total number of sites.
    While ideally we would estimate an overall value for $\theta$, it seems more
    important for the purposes of this investigation to estimate $\psi$.
